---
title: "Chocolate Ratings"
author: "Gary Lu"
date: "1/20/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

Ideas were taken from [Will Canniford at Kaggle](https://www.kaggle.com/willcanniford/chocolate-bar-ratings-extensive-eda) who supplied the dataset found on the [Tidy Tuesdays Github](https://github.com/rfordatascience/tidytuesday)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytuesdayR)
library(lubridate)
```

## Loading TidyTuesday Data

```{r, echo = FALSE}
tuesdata <- tidytuesdayR::tt_load('2022-01-18')

choco <- tuesdata$chocolate
```

```{r}
choco$review_date <- ymd(choco$review_date, truncated = 2L)

ratings <- choco %>% 
  group_by(review_date) %>% 
  summarise(med_rating = median(rating), n = n(), mean_rating = mean(rating), sd = sd(rating))

  ggplot(ratings, aes(x = review_date, y = med_rating))+
  geom_line()+
  geom_point()+
  labs(title = "Median rating of chocolates",
       x = "",
       y = "")+
  scale_x_continuous(breaks = sort(choco$review_date %>% unique()))
  
  ggplot(ratings, aes(x = review_date, y = mean_rating))+
  geom_line()+
  geom_point()+
  labs(title = "Mean rating of chocolates",
       x = "",
       y = "")+
  scale_x_continuous(breaks = sort(choco$review_date %>% unique()))
```

Median and means are around the same

```{r}
choco2 <- separate_rows(choco, c("most_memorable_characteristics"), sep = ",")
choco2$most_memorable_characteristics <- choco2$most_memorable_characteristics %>% str_trim()
```


```{r, echo=FALSE}
fourplus <- choco %>% filter(rating >= 4.0)

fourplus %>% 
  count(company_manufacturer, sort = TRUE) %>% 
  filter(n > 2) %>% 
  ggplot(aes(x = fct_reorder(company_manufacturer, -n), y = n))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  labs(title = "Manufacturers with at least 3 chocolates rated 4 or higher",
       x = "Manufacturer",
       y = "")
```

```{r}
# Soma specfic data since they have the most number of chocolates rated 4 or higher

soma <- choco %>% filter(company_manufacturer == "Soma")
soma2 <- choco2 %>% filter(company_manufacturer == "Soma")

soma_high_rated <- soma2 %>% filter(rating >= 4.0)

soma_high_rated %>% 
  group_by(review_date, most_memorable_characteristics) %>% 
  summarise(n = n()) %>% 
  ggplot()+
  geom_col(aes(x = most_memorable_characteristics, y = n))+
  facet_wrap(~review_date)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  labs(title = "Characteristics of high rated chocolate from Soma")
```


```{r}
choco_high_rated <- choco2 %>% filter(rating >= 4.0)

choco_high_rated %>% 
  group_by(most_memorable_characteristics) %>% 
  summarise(n = n()) %>% 
  filter(n > 1) %>% 
  ggplot()+
  geom_col(aes(x = fct_reorder(most_memorable_characteristics, -n), y = n))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  labs(title = "Characteristics of high rated chocolate (overall)",
       x = "",
       y = "")
```

Higher rated chocolates are described as either creamy, cocoa, or nutty


```{r}
# Group by country
choco %>% 
  group_by(review_date, company_location) %>% 
  filter(n() > 15) %>% 
  summarise(mean_rating = mean(rating)) %>% 
  ggplot()+
  geom_line(aes(x = review_date, y = mean_rating, color = company_location))+
  geom_point(aes(x = review_date, y = mean_rating, color = company_location))+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  labs(title = "Average ratings by country (at least 15 ratings)")+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{r}
# Ratings based on Cocoa percent
choco %>% 
  group_by(review_date, cocoa_percent) %>%
  filter(n() > 10) %>% 
  summarise(mean_rating = mean(rating)) %>% 
  ggplot(aes(group = 1))+
  geom_point(aes(review_date, y = mean_rating, color = cocoa_percent))+
  geom_line(aes(review_date, y = mean_rating, color = cocoa_percent))+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  labs(title = "Average ratings by cocoa percentage (at least 10 ratings)")+
  facet_wrap(~cocoa_percent)+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{r}
# Rating based on company

choco %>% 
  group_by(company_manufacturer) %>%
  filter(n() > 10) %>% 
  group_by(company_manufacturer, review_date) %>% 
  summarise(mean_rating = mean(rating)) %>% 
  ungroup() %>% 
  ggplot()+
  geom_point(aes(review_date, y = mean_rating, color = company_manufacturer))+
  geom_line(aes(review_date, y = mean_rating, color = company_manufacturer))+
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")+
  labs(title = "Average ratings by company (at least 10 ratings)")+
  facet_wrap(~company_manufacturer)+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

Possible factors that can contribute: low sample size, people may be biased if they know the brand they are reviewing

```{r}
# Ratings based on cocoa bean origin

choco %>% 
  group_by(country_of_bean_origin) %>%
  filter(n() > 10) %>% 
  group_by(country_of_bean_origin, review_date) %>% 
  summarise(mean_rating = mean(rating)) %>% 
  ungroup() %>% 
  ggplot()+
  geom_point(aes(review_date, y = mean_rating, color = country_of_bean_origin))+
  geom_line(aes(review_date, y = mean_rating, color = country_of_bean_origin))+
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")+
  labs(title = "Average ratings by country of bean origin (at least 10 ratings)")+
  facet_wrap(~country_of_bean_origin)+
  theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{r}
# Changes in rating statistics
choco %>% 
  group_by(review_date) %>% 
  summarise(n = n(), mean = mean(rating), median = median(rating), sd = sd(rating)) %>% 
  ggplot(aes(x = review_date))+
  geom_line(aes(y = mean, color = "mean"))+
  geom_line(aes(y = median, color = "median"))+
  labs(title = "Changes in mean and median rating each year")+
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")

choco %>% 
  group_by(review_date) %>% 
  summarise(n = n(), mean = mean(rating), median = median(rating), sd = sd(rating)) %>% 
  ggplot(aes(x = review_date, y = sd))+
  geom_point(aes(size = n))+
  geom_line()+
  labs(title = "Changes in sd",
       y = "Standard Deviation of ratings")+
  scale_x_date(date_breaks = "2 years", date_labels = "%Y")
```

The number of ratings goes up each year while the standard deviation decreases
The mean/median never falls under 3.0

```{r}
# Under 3.0 ratings

under3 <- choco %>% 
  group_by(review_date) %>% 
  filter(rating < 3.0)

under3 %>% 
  group_by(review_date) %>% 
  mutate(n = n(), mean = mean(rating)) %>% 
  ggplot(aes(x = review_date, y = rating))+
  geom_point()+
  geom_line(aes(y = mean, color = "Mean rating"))+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")
```

Over the years, the number of ratings < 3.0 have decreased as well as the overall spread of ratings