---
title: "Dog Breeds"
date: 2022-02-022
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(tidytuesdayR)
library(ggimage)
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

breed_traits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_traits.csv')
trait_description <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/trait_description.csv')
breed_rank_all <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_rank.csv')
```


# Readme

### The idea behind this analysis is to, based on the ratings, find ideal dog breeds

```{r}
trait_description
```
The Ideal dog breed should have high ratings in the "positive" categories. 

Some are subjective depending on a person's lifestyle and preference but the ones I will rate on will be

* Affectionate With Family

* Good With Young Children

* Good With Other Dogs

* Openness To Strangers

* Watchdog/Protective Nature

* Trainability Level

```{r}
positive_traits <-  c("Affectionate With Family", "Good With Young Children", "Good With Other Dogs", "Openness To Strangers", "Watchdog/Protective Nature", "Trainability Level")

df <- breed_traits %>% select(Breed, positive_traits)
df2 <- df %>% 
  pivot_longer(cols = 2:7, names_to = "Trait", values_to = "value") %>% 
  group_by(Breed)%>% 
  summarise(sum_traits = sum(value)) %>% 
  arrange(desc(sum_traits)) %>% 
  head(10)

df2$Breed <- df2$Breed %>% str_remove_all("[(|)]") %>% str_replace_all("\\s", "_")

df3 <- breed_rank_all %>% select(Breed, Image)
df3$Breed <- df3$Breed %>% str_remove_all("[(|)]") %>% str_replace_all("\\s", "_")

df3 <- df2 %>% left_join(df3)
df3$Breed <- df3$Breed %>% str_replace_all("_", " ")
```

```{r}
plot <- ggplot(df3, aes(x = fct_reorder(Breed, sum_traits), sum_traits, fill = Breed))+
  geom_col()+
  labs(title = "Highest Rated Dog Breeds",
       subtitle = "Based on Affection, Openness to Strangers, Protective Nature, Trainability",
       x = "",
       y = "Sum of Rating")+
  theme(legend.position = "")+
  coord_flip()

plot + geom_image(aes(image = Image), size = 0.09)
```

