---
title: "Thomas Mock (2021). Tidy Tuesday: A weekly data project aimed at the R ecosystem. https://github.com/rfordatascience/tidytuesday."
author: "Gary Lu"
date: "6/10/2021"
output: html_document
---

I will be using the 2021-04-20 dataset Netflix Titles at <https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-04-20/readme.md>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(stringi)
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv')
```

```{r}
# Helper Functions I will probably use a lot

group_count <- function(df, column_names){
  df %>% 
    group_by_at(column_names) %>% 
    summarise(n = n()) %>% 
    arrange(desc(n))
}

row_search <- function(df, col, string){
  df %>% 
    filter(str_detect(df[[col]], str_c("^", string)))
}
```

```{r, Breakdown of media types on Netflix}

netflix_titles %>% group_count("type") %>% 
  mutate(percent = n / sum (n) * 100) %>% 
  ggplot()+
  geom_col(aes(x = type, y = percent, fill = type))
```

Using helper function group_count can be used to see the most popular of any column

```{r}
netflix_titles$date_added <- mdy(netflix_titles$date_added)
netflix_countries_raw <- netflix_titles %>% group_count("country")
netflix_countries_raw
```

Multiple countries exist, I split them into separate rows by country

```{r}
netflix2 <- separate_rows(netflix_titles, country, sep = ", ")

netflix2 %>% 
  group_count("country") %>% 
  head(20) %>% 
  ggplot(aes(x = fct_reorder(country, n), y = n))+
  geom_col()+
  coord_flip()+
  labs(title = "Country where the movie / show was produced",
      x = "country")
```

Some series has NA in the country, which is weird because everything has to be filmed SOMEWHERE right?

```{r}
na_country <- netflix2 %>% filter(is.na(country))
na_country
```

Using separate_rows and group_count for directors
```{r}
separate_rows(netflix_titles, director, sep = ", ") %>%  
  group_count("director")
na_directors <- netflix_titles %>% filter(is.na(director))
```

a LOT of missing info for directors (and the dataset as a whole) for some reason

EDA while keeping NA values:

```{r}
# data set that only contains at least 1 NA
netflix_titles_na <- netflix_titles %>% filter(!complete.cases(netflix_titles))
colSums(is.na(netflix_titles_na))
```

Here we see how the NA values are spread out


Using data with NA filtered out
```{r}
netflix2 %>% filter(complete.cases(netflix2)) %>% 
  group_count("country") %>% 
  head(20) %>% 
 ggplot()+
   geom_col(aes(x = fct_reorder(country, n), y = n))+
   coord_flip()+
   labs(title = "Country where the movie / show was produced",
       x = "country",
       subtitle = "Missing values removed")
```

Interesting takeaways:
Canada jumps to the 4th spot after removing any row with missing values 
Japan and South Korea which were the previous 4th and 5th most popular country has dropped

```{r, South Korean and Japan titles}
netflix_JPSK <- netflix_titles %>% 
                filter(country == "Japan" | country == "South Korea") %>% 
                filter(!complete.cases(.))

colSums(is.na(netflix_JPSK))
```

For Japanese and South Korean titles, most missing values are in the director column.

```{r}
netflix_titles %>% 
  separate_rows(listed_in, sep = ", ") %>% 
  group_count("listed_in") %>% 
  head(20) %>% 
  ggplot()+
  geom_col(aes(x = fct_reorder(listed_in, n), y = n))+
  coord_flip()+
  labs(title = "Top 20 Most popular Genres",
      x = "Genre")
```


```{r}
netflix_titles %>% 
  group_count("release_year") %>% 
  head(20) %>% 
  ggplot()+
  geom_col(aes(x = release_year, y = n))+
  scale_x_continuous(n.breaks = 20)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  labs(title = "Number of titles released per year")
```

```{r}
timediff <- netflix_titles %>% 
            mutate(timediff = year(date_added) - release_year) %>% 
            select(title, date_added, release_year, timediff)

ggplot(timediff, aes(x = release_year, y = timediff, color = release_year))+
  geom_point()+
  labs(title = "Time it takes for something to get added to Netflix",
       y = "Days to get added to Netflix")
```

```{r}
x <- netflix_titles$description %>% 
  str_split(boundary("word")) %>% 
  unlist() %>% 
  str_extract("^.{5,}") %>% 
  as_tibble()

x <- x %>%  filter(!is.na(x))

rename(x, word = value) %>% 
  group_count("word") %>% 
  head(15) %>% 
  ggplot()+
  geom_col(aes(x = fct_reorder(word, n), y = n, fill = word))+
  coord_flip()+
  labs(title = "15 Most common words in description (at least 5 letters)",
       x = "Word")+
  theme(legend.position = "")
```
```{r, Country specific functions}

countrydata <- function(df, input){
 
  newdf <- df %>% 
            filter(country == input)
  
top_genres <- newdf %>% 
              separate_rows(listed_in, sep = ", ") %>% 
              group_count("listed_in") %>% 
              head(15) %>% 
              ggplot()+
              geom_col(aes(x = fct_reorder(listed_in, n), y = n))+
              coord_flip()+
              labs(title = "Top 20 Most Popular Genres based on film location",
                    x = "Genre",
                   subtitle = input)


top_directors <- newdf %>%
              filter(!is.na(newdf$director)) %>% 
              separate_rows(director, sep = ", ") %>% 
              group_count("director") %>% 
              head(15) %>% 
              ggplot()+
              geom_col(aes(x = fct_reorder(director, n), y = n))+
              coord_flip()+
              labs(title = "Top directors",
                    x = "director",
                   subtitle = input)

print(top_genres)
print(top_directors)

}

```

An example of the above function:
```{r}
netflix_titles %>% 
  countrydata("United States")

netflix_titles %>% 
  countrydata("Canada")
```